Notes on the Godunov Step in {\tt MAESTRO} and {\tt VARDEN}

### 2008-04-18 docs/Godunov_notes/paper.tex
added notes on conservative vs non-conservative s^edge prediction in varden

The extrapolation of $u$ to r-faces is completely analogous to the x-faces:

@@

The full-dimensional extrapolation of $s$ to r-faces is completely analogous to the x-faces:

@@

The normal velocity Riemann problem is:

@@

Do the exact analogous thing to get both components of velocity at the y-face, $\Ub_{\ib-\half\eb_y}^{1D}$. \\

Now we do a full-dimensional extrapolation to get only the normal velocity at each face:

@@

Then, solve a Riemann problem:

@@

Use the exact analogous procedure to obtain $v^{\mac,*}_{\ib-\half\eb_y}$.

@@

This is more complicated than the 2D case because we include corner coupling.  We compute $\Ub_{\ib-\half\eb_x}, \Ub_{\ib-\half\eb_y}$, and $\Ub_{\ib-\half\eb_z}$ in an analogous manner as equations (\ref{varden U_L^1D})-(\ref{Transverse Velocity Riemann Problem}).  Then, we compute an intermediate state, $u_{\ib-\half\eb_y}^{y|z}$, which is described as ``state $u_{\ib-\half\eb_y}^{1D}$ that has been updated to account for the transverse fluxes in the z direction'', using:

@@

&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_z}^{1D}+v_{\ib-\half\eb_z}^{1D}\right)\left(u_{\ib+\half\eb_z}^{z|y}-u_{\ib-\half\eb_z}^{z|y}\right).

@@

To compute $\Ub^{\edge}$, {\tt VARDEN} uses the exact same algorithm as the $s^{\edge}$ case in {\tt MAESTRO}.  To the algorithm for $\rho^{\edge}$ in {\tt VARDEN} is slightly different than the $s^{\edge}$ case in {\tt MAESTRO} since it uses a ``conservative'' formulation.  

@@

\rho_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib-\eb_x}^n, \label{varden rho_L^1D}\\
\rho_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib}^n. \label{varden rho_R^1D}

@@

\rho_{\ib-\half\eb_x}^{1D} =

@@

\half\left(\rho_{L,\ib-\half\eb_x}^{1D} + \rho_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
\rho_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
\rho_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\

@@

The procedure to obtain $\rho_{\ib-\half\eb_y}^{1D}$ is completely analogous.  Now we do a full-dimensional extrapolation of $\rho$ to each face.  The extrapolation of $\rho$ to x-faces is:

@@

\rho_{L,\ib-\half\eb_x}^{\edge} &=& \rho_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib-\eb_x} - (\rho\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} \nonumber \\
&&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right), \nonumber \\

@@

\rho_{R,\ib-\half\eb_x}^{\edge} &=& \rho_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib} - (\rho\nabla\cdot\Ub^{\mac})_{\ib} \nonumber \\
&&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right).

@@

\rho_{\ib-\half\eb_x}^{\edge} =

@@

\half\left(\rho_{L,\ib-\half\eb_x}^{\edge} + \rho_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
\rho_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
\rho_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.

@@

The procedure to obtain $\rho_{\ib-\half\eb_y}^{\edge}$ is completely analogous.

@@

This is more complicated than the 2D case because we include corner coupling.  We compute $\rho_{\ib-\half\eb_x}^{1D}$, $\rho_{\ib-\half\eb_y}^{1D}$, and $\rho_{\ib-\half\eb_z}^{1D}$ in a completely analogous manner to equations (\ref{varden rho_L^1D}) and (\ref{varden rho_R^1D}).

>>>>

The extrapolation of $u$ to r-faces is analogous to the extrapolation to x-faces:

@@

The full-dimensional extrapolation of $s$ to r-faces is analogous to the extrapolation to x-faces:

@@

We obtain the normal velocity using the Riemann problem:

@@

We perform analogous operations to compute both components of velocity at the y-faces, $\Ub_{\ib-\half\eb_y}^{1D}$. \\

Now we do a full-dimensional extrapolation to get the MAC velocity at the x-faces (note that we only compute the normal components):

@@

Then, we solve a Riemann problem:

@@

We perform analogous operations to compute the normal velocity at the y-faces, $v^{\mac,*}_{\ib-\half\eb_y}$.

@@

This is more complicated than the 2D case because we include corner coupling.  We compute $\Ub_{\ib-\half\eb_x}^{1D}, \Ub_{\ib-\half\eb_y}^{1D}$, and $\Ub_{\ib-\half\eb_z}^{1D}$ in an analogous manner as equations (\ref{varden U_L^1D})-(\ref{Transverse Velocity Riemann Problem}).  Then, we compute an intermediate state, $u_{\ib-\half\eb_y}^{y|z}$, which is described as ``state $u_{\ib-\half\eb_y}^{1D}$ that has been updated to account for the transverse derivatives in the z direction'', using:

@@

&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_z}^{1D}+w_{\ib-\half\eb_z}^{1D}\right)\left(u_{\ib+\half\eb_z}^{z|y}-u_{\ib-\half\eb_z}^{z|y}\right).

@@

To compute $\Ub^{\edge}$, {\tt VARDEN} uses the exact same algorithm as the $s^{\edge}$ case in {\tt MAESTRO}.  The algorithm for $\rho^{\edge}$ in {\tt VARDEN} is slightly different than the $s^{\edge}$ case in {\tt MAESTRO} since it uses a ``conservative'' formulation.  Here, $s$ is used in place of either $\rho, u, v$, or $w$ (in 3D).

@@

s_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib-\eb_x}^n, \label{varden s_L^1D}\\
s_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib}^n. \label{varden s_R^1D}

@@

@@

@@

We use an analogous procedure to obtain $s_{\ib-\half\eb_y}^{1D}$.  Now we do a full-dimensional extrapolation of $s$ to each face.  The extrapolation of a ``conserved'' $s$ to x-faces is:

@@

s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib-\eb_x} - (s\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} \nonumber \\
&&- \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] + \frac{\dt}{2h}s_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right), \nonumber \\

@@

s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib} - (s\nabla\cdot\Ub^{\mac})_{\ib} \nonumber \\
&&- \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\half\eb_y}\right] + \frac{\dt}{2h}s_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right).
The extrapolation of a ``non-conserved'' $s$ to x-faces is:
s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib-\eb_x} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}+v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{1D} - s_{\ib-\eb_x-\half\eb_y}^{1D}\right), \nonumber \\
s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib} - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\mac}+v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{1D} - s_{\ib-\half\eb_y}^{1D}\right).

@@

@@

@@

We use an analogous procedure to compute $s_{\ib-\half\eb_y}^{\edge}$.

@@

This is more complicated than the 2D case because we include corner coupling.  We compute $s_{\ib-\half\eb_x}^{1D}$, $s_{\ib-\half\eb_y}^{1D}$, and $s_{\ib-\half\eb_z}^{1D}$ in an analogous manner to equations (\ref{varden s_L^1D}) and (\ref{varden s_R^1D}).
